---
title: "ShovanBiswas-DATA24_Homework_10"
author: "Shovan Biswas"
date: "11/30/2020"
output:
  pdf_document: default
  word_document: default
  html_document:
    highlight: pygments
    theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# knitr::opts_chunk$set(fig.width = 8, fig.height = 8) 
```

# Libraries  

```{r, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
library(corrplot)
library(reshape2)
library(Amelia)
library(dlookr)
library(fpp2)
library(plotly)
library(gridExtra)
library(readxl)
library(ggplot2)
library(urca)
library(tseries)
library(AppliedPredictiveModeling)
library(RANN)
library(psych)
library(e1071)
library(corrplot)
library(glmnet)
library(mlbench)
library(caret)
library(earth)
library(randomForest)
library(party)
library(Cubist)
library(gbm)
library(rpart)
library(dplyr)
library(arulesViz)
library(igraph)
```

# Problem statement      

Imagine 10000 receipts sitting on your table. Each receipt represents a transaction with items that were purchased. The receipt is a representation of stuff that went into a customer’s basket - and therefore "Market Basket Analysis".     

That is exactly what the Groceries Data Set contains: a collection of receipts with each line representing 1 receipt and the items purchased. Each line is called a transaction and each column in a row represents an item.  The data set is attached.      

Your assignment is to use R to mine the data for association rules. You should report support, confidence and lift and your top 10 rules by lift.       

Extra credit: do a simple cluster analysis on the data as well.  Use whichever packages you like.  Due May 3 before midnight.      


# Brief explanation               

Initially, I proceeded to read with read_csv. Although I was able to read the usual csv file (GroceryDataSet.csv i.e.), it didn't help in down stream analysis. So, in order to mine the data for **Association Rules**, I googled and learned that apriori() function was required. This is not something, which we customarily use or have used so in the past. On googling, I hit upon the following page:        

    https://blog.aptitive.com/building-the-transactions-class-for-association-rule-mining-in-r-using-arules-and-apriori-c6be64268bc4                     

The page gives an overview of transactions class, apriori() functions etc. The package arules is required, which I added to the list of libraries above. "Market Basket Analysis" was a good clue.          
Explanation of some of the terms in Association Rules, which we'll encounter below:          

Support of a set of items is the frequency with which, an item appears in the dataset.        

Confidence of a rule is the frequency of how often a rule has been found to be true.       

Lift is the ratio of the actual support to the expected support.        

# Reading data and summary               


```{r}
# grocery_transactions <- read_csv('./GroceryDataSet.csv')
grocery_transactions <- read.transactions('./GroceryDataSet.csv', sep = ",")
```

```{r}
summary(grocery_transactions)
```

From summary, we see that some of the most freuent items are "whole milk", "other vegetables", "rolls/buns", "soda" etc. In order to get a better visualization, I'll use function itemFrequencyPlot().      

# Frequency of top 20 most frequent items                

```{r}
itemFrequencyPlot(grocery_transactions, topN = 20, type = "absolute", main = "Top 20 Items")
```

This graph gives an idea of frequencies of top 20 most frequent items. This graph corroborate the few observations in summary.                


# Further analysis                              

Now, I'll use apriori() function, for "Market Basket Analysis". I explored apriori() function, by varying the values of the parameters, **support** and **confidence**. With some combinations, I didn't get any results at all -- simply errored out. With support = 0.001, confidence = 0.4, in descending order of lift, I got a table (shown down below).       

```{r}
support <- 0.001
confidence <- 0.4
rules <- apriori(grocery_transactions, parameter = list(support = support, confidence = confidence), control = list(verbose = FALSE))
```

```{r}
summary(rules)
```

An important observation in summary is, there 8955 rules with length from 2 to 6.       

In the following, I'll display the top 10 rules with their support and confidence, sorted descending order of lift.        

```{r}
rules %>% DATAFRAME() %>% arrange(desc(lift)) %>% top_n(10) %>% kable()
```

What is this table telling us? The rule having the greatest lift (2.246605), is for the item **{other vegetables}**, after purchase of **{root vegetables}**. The support and confidence of the item are 0.04738180 and 0.4347015 respectively.        


The following graph gives a good visualization of how the items are associating.           

```{r}
subrules <- head(rules, n = 10, by = 'lift')
plot(subrules, method = 'graph')
```

# Cluster analysis                      

In order to do cluster analysis, groupings must be identified. After creating a network graph from the given data, I'll use cluster_louvain() to 

```{r}
grocery_csv <- read.csv("GroceryDataSet.csv", header = FALSE) %>% mutate(shoper_id = row_number()) %>% pivot_longer(-shoper_id) %>% filter(value != "") %>% select(-name)

communities <- grocery_csv %>% rename(to = value, from = shoper_id) %>% graph_from_data_frame(directed = FALSE) %>% cluster_louvain() %>% communities()
```

The following step will associate customers and items to 19 clusters.       

```{r}
products <- as.character(unique(grocery_csv$value))

df <- data.frame(name = c(NA), members = c(NA)) %>% na.omit() # create data frame

for (i in 1:length(communities)){
  cluster_name <- paste0(i,": ")
  cluster_members <- 0
  for (member in communities[[i]]){
    if (member %in% products){
      cluster_name <- paste0(cluster_name, member, " + ")
    } else {
      cluster_members <- cluster_members + 1
    }
  }
  cluster_name <- substr(cluster_name,1,nchar(cluster_name)-3)
  df <- rbind(df, data.frame(name = cluster_name, members = cluster_members))
}
  df %>%
  arrange(desc(members)) %>% kable()
```

