---
title: "ShovanBiswas-DATA24_Homework_08"
author: "Shovan Biswas"
date: "10/30/2020"
output:
  pdf_document: default
  word_document: default
  html_document:
    highlight: pygments
    theme: cerulean
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 8, fig.height = 8) 
```

# Libraries  

```{r, eval=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(kableExtra)
library(corrplot)
library(reshape2)
library(Amelia)
library(dlookr)
library(fpp2)
library(plotly)
library(gridExtra)
library(readxl)
library(ggplot2)
library(urca)
library(tseries)
library(AppliedPredictiveModeling)
library(RANN)
library(psych)
library(e1071)
library(corrplot)
library(glmnet)
library(mlbench)
library(caret)
library(earth)
```

# Exercise 7.2  

Friedman (1991) introduced several benchmark data sets create by simulation. One of these simulations used the following nonlinear equation to create data:      

$y = 10 sin(\pi x_1 x_2) + 20(x_3 - 0.5)^2 + 10x_4 + 5x_5 + N(0, \sigma^2)$         

where the $x$ values are random variables uniformly distributed between $[0, 1]$ (there are also $5$ other non-informative variables also created in the simulation). The package **mlbench** contains a function called mlbench.friedman1 that simulates these data:      

(This exercise is based on library(mlbench), which I included in libraries at the top.)     
```{r}
set.seed(200)

trainingData <- mlbench.friedman1(200, sd = 1)

## We convert the 'x' data from a matrix to a data frame
## One reason is that this will give the columns names.

trainingData$x <- data.frame(trainingData$x)

## Look at the data using

featurePlot(trainingData$x, trainingData$y)

## or other methods.
```


```{r}
## This creates a list with a vector 'y' and a matrix
## of predictors 'x'. Also simulate a large test set to
## estimate the true error rate with good precision:

testData <- mlbench.friedman1(5000, sd = 1)
testData$x <- data.frame(testData$x)
```

Tune several models on these data. For example:    

(I included library caret in libraries at the top.)     
```{r}
knnModel <- train(x = trainingData$x, y = trainingData$y, method = "knn", preProc = c("center", "scale"), tuneLength = 10)
knnModel
```

```{r}
knnPred <- predict(knnModel, newdata = testData$x)

## The function 'postResample' can be used to get the test set
## perforamnce values
postResample(pred = knnPred, obs = testData$y)
```

Which models appear to give the best performance? Does MARS select the informative predictors (those named X1–X5)?        


## Answer:      

We observed above that the RMSE of kNN model is 3.2932153. In the following, we'll explore all other models, Neural Networks, MARS and SVM, mentioned in the book, in this order.     

### Neural Networks      

Used code from page 163 of textbook.      
```{r}
nnetGrid <- expand.grid(.decay = c(0, 0.01, .1), .size = c(1:10), .bag = FALSE)

nnetTune <- train(x = trainingData$x, y = trainingData$y, method = "avNNet", tuneGrid = nnetGrid,
                  trControl = trainControl(method = "cv"), preProc = c("center", "scale"),
                  linout = TRUE, trace = FALSE,
                  MaxNWts = 10 * (ncol(trainingData$x) + 1) + 10 + 1,
                  maxit = 500)

nnetTune
```

```{r}
Neural_pred <- postResample(pred = predict(nnetTune, newdata = testData$x), obs = testData$y)
Neural_pred
```

Observation: RMSE of Neural Networks is 2.496722. It's way higher than what we obtained in kNN (3.2040595).      

```{r}
varImp(nnetTune)
```

The top 5 variables are X4, X1, X2, X5, X3.       



### MARS      

Used code from page 165 of textbook. Included library(earth) in libraries at the top.         
```{r}
marsGrid <- expand.grid(.degree = 1:2, .nprune = 2:38)

marsTuned <- train(x = trainingData$x, y = trainingData$y, method = "earth", tuneGrid = marsGrid,
                   trControl = trainControl(method = "cv"))

marsTuned
```

```{r}
MARS_pred <- postResample(pred = predict(marsTuned, newdata = testData$x), obs = testData$y)
MARS_pred
```

Observation: RMSE of MARS is 1.2803060. It's least and best so far.      

```{r}
varImp(marsTuned)
```

The top 5 variables are X1, X4, X2, X5, X3.       




### Support Vector Machines      

Used code from page 167 of textbook.      
```{r}
svmRTuned <- train(x = trainingData$x, y = trainingData$y,
                   method = "svmRadial", preProc = c("center", "scale"),
                   tuneLength = 14, trControl = trainControl(method = "cv"))

svmRTuned
```

```{r}
SVM_pred <- postResample(pred = predict(svmRTuned, newdata = testData$x), obs = testData$y)
SVM_pred
```

Observation: RMSE of SVM is 2.0469184.            

```{r}
varImp(svmRTuned)
```

The top 5 variables are X4, X1, X2, X5, X3.       



## Summary      

```{r}
results <- data.frame(t(postResample(pred = knnPred, obs = testData$y))) %>% mutate("Model" = "KNN")

results <- data.frame(t(Neural_pred)) %>% mutate("Model"= "Neural Networks") %>% bind_rows(results)

results <- data.frame(t(MARS_pred)) %>% mutate("Model"= "MARS") %>% bind_rows(results)

results <- data.frame(t(SVM_pred)) %>% mutate("Model"= "Support Vector Machines") %>% bind_rows(results)
```

```{r}
results %>% select(Model, RMSE, Rsquared, MAE) %>% arrange(RMSE)
```

## Conclusion

MARS outperformed the others.      


# Exercise 7.5      

Exercise 6.3 describes data for a chemical manufacturing process. Use the same data imputation, data splitting, and pre-processing steps as before and train several nonlinear regression models.     

## Answer:      

Before we begin answring the questions, let's pre-process the data. In previous Homework 7, this data had to be imputed and Box-coxed. I'll do the same this time, but will not inspect with histograms, because we already know about that.        

```{r}
data(ChemicalManufacturingProcess)
dim(ChemicalManufacturingProcess)
```

I'll impute the data using the same technique i.e. mice().    

```{r}
data_imputed <- mice(ChemicalManufacturingProcess, m = 1, method = "pmm", print = F) %>% complete()
```

```{r}
any(is.na(data_imputed))
```

At this point, the data is imputed. I'll proceed to Box-Cox it.     

```{r}
# Initially identifying columns, whose skewness are not less than 1.
transform_cols <- c()

for(i in seq(from = 1, to = length(data_imputed), by = 1)) {
  if(abs(skewness(data_imputed[, i])) >= 1) {
    transform_cols <- append(transform_cols, names(data_imputed[i]))
  }
}

# Applying Box-cox.
lambda <- NULL
data_imputed_2 <- data_imputed

for (i in 1:length(transform_cols)) {
  lambda[transform_cols[i]] <- BoxCox.lambda(abs(data_imputed[, transform_cols[i]]))

  data_imputed_2[c(transform_cols[i])] <- BoxCox(data_imputed[transform_cols[i]], lambda[transform_cols[i]])
}
```

At this point, the data is pre-processed. The pre-processed data is stored in the variable data_imputed_2.     

So, I'll proceed to split the data into train and test in 80/20 ratio.      

```{r}
set.seed(200)

split_index <- createDataPartition(data_imputed_2$Yield, p = 0.8, list = FALSE)
X_train <- data_imputed_2[split_index, ]
y_train <- data_imputed_2$Yield[split_index]

X_test <- data_imputed_2[-split_index, ]
y_test <- data_imputed_2$Yield[-split_index]
```




## (a)Which nonlinear regression model gives the optimal resampling and test set performance?        

### kNN      

```{r, eval=TRUE, message=FALSE, warning=FALSE}
knnModel <- train(x = X_train, y = y_train, method = "knn", preProc = c("center", "scale"), tuneLength = 10)
knnModel
```

```{r}
knnPred <- postResample(pred = predict(knnModel, newdata = X_test), obs = y_test)
knnPred
```

### Neural Networks      

```{r}
set.seed(200)

nnetGrid <- expand.grid(.decay = c(0, 0.01, .1), .size = c(1:10), .bag = FALSE)

nnetTune <- train(x = X_train,
                  y = y_train,
                  method = "avNNet",
                  tuneGrid = nnetGrid,
                  trControl = trainControl(method = "cv"),
                  preProc = c("center", "scale"),
                  linout = TRUE,
                  trace = FALSE,
                  MaxNWts = 10 * (ncol(X_train) + 1) + 10 + 1,
                  maxit = 500)

nnetTune
```


```{r}
Neural_pred <- postResample(pred = predict(nnetTune, newdata = X_test), obs = y_test)
Neural_pred
```

### MARS      

```{r}
marsGrid <- expand.grid(.degree = 1:2, .nprune = 2:38)

marsTuned <- train(x = X_train, y = y_train, method = "earth", tuneGrid = marsGrid, preProc = c("center", "scale"), trControl = trainControl(method = "cv"))

marsTuned
```

```{r}
MARS_pred <- postResample(pred = predict(marsTuned, newdata = X_test), obs = y_test)
MARS_pred
```

### SVM

```{r, eval=TRUE, message=FALSE, warning=FALSE}
svmRTuned <- train(x = X_train, y = y_train,
                   method = "svmRadial", preProc = c("center", "scale"),
                   tuneLength = 14, trControl = trainControl(method = "cv"))

svmRTuned
```

```{r}
SVM_pred <- postResample(pred = predict(svmRTuned, newdata = X_test), obs = y_test)
SVM_pred
```

## Summary      

```{r}
results <- data.frame(t(knnPred)) %>% mutate("Model" = "KNN")

results <- data.frame(t(Neural_pred)) %>% mutate("Model"= "Neural Networks") %>% bind_rows(results)

results <- data.frame(t(MARS_pred)) %>% mutate("Model"= "MARS") %>% bind_rows(results)

results <- data.frame(t(SVM_pred)) %>% mutate("Model"= "Support Vector Machines") %>% bind_rows(results)
```

```{r}
results %>% select(Model, RMSE, Rsquared, MAE) %>% arrange(RMSE)
```

## Conclusion

MARS outperformed the others.      


(b) Which predictors are most important in the optimal nonlinear regression model? Do either the biological or process variables dominate the list? How do the top ten important predictors compare to the top ten predictors from the optimal linear model?       

```{r}
varImp(marsTuned)
```

varImp(marsTuned) did not return any predictor. So, let me try the second best mode i.e. Neural Networks.

```{r}
varImp(nnetTune)
```

In the case of Neural Networks, 6 of the top ten predictors are ManufacturingProcess predictors and 3 are BiologicalMaterial. So, the ManufacturingProcess predictors dominate.     


## PLS_MODEL        
```{r}
set.seed(200)
pls_model <- train(x = X_train, y = y_train, method = "pls", preProc = c("center", "scale"), trControl = trainControl("cv", number = 10), tuneLength = 25)

pls_prediction <- predict(pls_model, newdata = X_test)

results <- data.frame(Model = "PLS", RMSE = caret::RMSE(pls_prediction, y_test), Rsquared = caret::R2(pls_prediction, y_test))
results
```


```{r}
varImp(pls_model)
```

In nonlinear models, MARS performed best. The RMSE was very close to zero. But PLS_MODEL, which is linear, returned an RMSE of 0.1308365 is higher than MARS's RMSE. So, it faired worse.       

However, in linear model PLS_MODEL, among the top 10 variables, ManufacturingProcess is dominant.      


(c) Explore the relationships between the top predictors and the response for the predictors that are unique to the optimal nonlinear regression model. Do these plots reveal intuition about the biological or process predictors and their relationship with yield?       

By comparing 20 varImp(nnetTune) with 20 varImp(pls_model), the former being nonlinear and latter liner, we get the following predictors as unique to nonlinear.      

BiologicalMaterial03
BiologicalMaterial09
ManufacturingProcess11
ManufacturingProcess30

This was done offline manually on Bash-shell.

The plots of each of these variables that are unique to nonlinear set are shown below.     


```{r}
ggplot(X_train, aes(BiologicalMaterial03, Yield)) + geom_point() + ggtitle("BiologicalMaterial03")
```

```{r}
ggplot(X_train, aes(ManufacturingProcess34, Yield)) + geom_point() + ggtitle("BiologicalMaterial09")
```


```{r}
ggplot(X_train, aes(ManufacturingProcess34, Yield)) + geom_point() + ggtitle("ManufacturingProcess11")
```

```{r}
ggplot(X_train, aes(ManufacturingProcess34, Yield)) + geom_point() + ggtitle("ManufacturingProcess30")
```

These don't indicate any special relationship.      